---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```
# redivis.bigrquery

The `redivis.bigrquery` package is a light fork of the awesome [bigrquery](https://github.com/r-dbi/bigrquery) package, allowing you to leverage its functionality
to interface with tables stored on Redivis. All authentication and billing is managed via your Redivis API credentials.

Please note that the only supported methods are those that involve querying tables. Interfaces involved in listing BigQuery resource, reference BigQuery datasets, or any calls to create, modify, or delete BigQuery resources are not supported.

### Installation

Install from GitHub:

```R
# install.packages('devtools')
devtools::install_github("redivis/bigrquery")
```

### [bigrquery](https://github.com/r-dbi/bigrquery)

The bigrquery package makes it easy to work with data stored in
[Google BigQuery](https://developers.google.com/bigquery/) by allowing you to query BigQuery tables and retrieve metadata about your projects, datasets, tables, and jobs. The bigrquery package provides three levels of abstraction on top of BigQuery:

* The low-level API provides thin wrappers over the underlying REST API. All
  the low-level functions start with `bq_`, and mostly have the form
  `bq_noun_verb()`. This level of abstraction is most appropriate if you're
  familiar with the REST API and you want do something not supported in the
  higher-level APIs.

* The [DBI interface](http://www.r-dbi.org) wraps the low-level API and
  makes working with BigQuery like working with any other database system.
  This is most convenient layer if you want to execute SQL queries in
  BigQuery or upload smaller amounts (i.e. <100 MB) of data.

* The [dplyr interface](http://dbplyr.tidyverse.org/) lets you treat BigQuery
  tables as if they are in-memory data frames. This is the most convenient
  layer if you don't want to write SQL, but instead want dbplyr to write it
  for you.

## Usage

### Low-level API

```{r}
library(redivis.bigrquery)
sql <- "SELECT * FROM `stanfordphs.commuting_zone_life_expectancy_trends.v1_0` LIMIT 10"

tb <- bq_project_query(sql)
bq_table_download(tb, max_results = 10)
```

### DBI

```{r, warning = FALSE}
library(DBI)

con <- dbConnect(redivis.bigrquery::bigquery())
sql <- "SELECT * FROM `stanfordphs.commuting_zone_life_expectancy_trends.v1_0` LIMIT 10"

dbGetQuery(con, sql, n = 10)
```

### dplyr

```{r, message = FALSE}
library(DBI)
library(dplyr)

con <- dbConnect(redivis.bigrquery::bigquery())

providers <- tbl(con, "ianmathews91.medicare_public_example.high_cost_in_providers_in_CA_output")

providers %>%
  select(drg_definition, provider_id, provider_state, mean_drg_cost) %>%
  head(10) %>%
  collect()
```
## Referencing tables
All tables belong to either a dataset or project. Table references use the following pattern:
```sql
# Datasets
# Note: table_name for datasets should always be "main". 
# An upcoming release of Redivis will allow for named tables within datasets.
owner_name.dataset_name[:v{version}][:sample][:{dataset_id}].table_name[:{table_id}]

# Projects
owner_name.project_name[:{project_id}].table_name[:{table_id}]
```
Note that datasets (and their tables) that are added to a project should still be referenced via their dataset identifier; the tables are not considered to be within the project. (`TODO: reconsider?`)

All non alpha-numeric and underscore characters in names and version tags can be replaced by an underscore (`_`) character. Colons (`:`), periods (`.`), and backticks (`` ` ``) must be replaced. Multiple underscores can be collapsed into a single underscore, and leading and trailing underscores can be ignored.

References can include several optional additions, denoted by a `:`. The `:sample` suffix on a dataset will reference that dataset's sample. The `:v{version}` suffix references a particular version of a dataset; if no version is specified, the latest version will be used. The `:{datasetId}`, `:{tableId}`, and `:{projectId}` suffixes allow you to specify the relevant persistent identifier, which will be used in place of the name and avoid future errors if table or dataset names are modified. 

If the `:tableId` is specified, no owner or project|dataset identifier is required. If the `:datasetId|:projectId` is specified, no owner is required.

### Examples

We can reference the [IPUMS 1940 Households](https://redivis.com/StanfordPHS/datasets/152) dataset as:
```sql
SELECT [] FROM `stanfordphs.IPUMS 1940 Households.main` 
```
We can reference without backticks if we escape non-word characters. Note that references are case-insensitive.
```sql
SELECT [] FROM stanfordphs.ipums_1940_households.main
```
By default this uses the lates (2.0) version of the dataset. If we want to work with version 1.0:
```sql
SELECT [] FROM stanfordphs.ipums_1940_households:v1_0.main
```
If we want to work witht the 1% sample:
```sql
# The order of the suffixes does not matter
SELECT [] FROM stanfordphs.ipums_1940_households:v1_0:sample.main
```
Finally, we can provide persistent ids to prevent our reference from breaking if an item is renamed:
```sql
SELECT [] FROM stanfordphs.ipums_1940_households:152:v1_0:sample.main
# We don't actually need the owner if an id is provided
SELECT [] FROM ipums_1940_households:152:v1_0:sample.main
# Can also provide a table id
SELECT [] FROM ipums_1940_households:152.main:178
# Don't need the dataset name if tableId is provided
SELECT [] FROM main:178
# Or even the table name
SELECT [] FROM :178
```
Referencing tables in [a project](https://redivis.com/projects/1008/tables/9443) is quite similar, though projects don't have versions or samples:
```sql
SELECT [] FROM ianmathews91.medicare_public_example.high_cost_in_providers_in_CA_output
# or
SELECT [] FROM medicare_public_example:1008.high_cost_in_providers_in_CA_output
# or 
SELECT [] FROM high_cost_in_providers_in_CA_output:9443
```

## Important details

### Authentication and authorization

The `REDIVIS_API_TOKEN` environment variable must be set to your Redivis API token, and the token must have `data.data` scope. Go to your [workspace/settings](https://redivis.com/workspace/settings) to create an API token. For example:
```
REDIVIS_API_TOKEN=your_api_token R ...
# or
Sys.setenv(REDIVIS_API_TOKEN = "your_api_token")
```

## Useful links

* [SQL reference](https://developers.google.com/bigquery/query-reference)
* [API reference](https://developers.google.com/bigquery/docs/reference/v2/)
